

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)

setwd("~/Desktop/Independent Study/Data")
IS_Data <- read_excel("~/Desktop/Independent Study/Data/IS_Data.xlsx")
```

## Cleaning/Prep
```{r}
# Remove unnecessary variables
IS_Data <- IS_Data[, -c(14, 16:ncol(IS_Data))]

# Renaming variables with spaces
IS_Data <- IS_Data |>
  rename(
    Yield = `Yield (%)`,
    RSR = `RSR (%)`,
    GHI = GHI_kWh_m2
  )

#Fix error in inconsistent input of climate zone abbreviations
IS_Data <- IS_Data |>
  #Bsk -> BSk
  mutate(Climate_zone = ifelse(Climate_zone == 'Bsk', 'BSk', Climate_zone)) |>
  #Cwb -> Cwa
  mutate(Climate_zone = ifelse(Climate_zone == 'Cwb', 'Cwa', Climate_zone))

#Recode shade types
IS_Data <- IS_Data |>
  #"." -> "Not Specified"
  mutate(Shade_type = ifelse(Shade_type == '.', 'Not specified', Shade_type)) |>
  #Solar_panel" -> "Solar panel"
  mutate(Shade_type = ifelse(Shade_type == "Solar_panel", "Solar panel", Shade_type))

# Remove shade type that is not solar panel
SolarPanelShade <- IS_Data[IS_Data$Shade_type == "Solar panel", ]

# Remove intercropping experiments
ArtificialShade <- IS_Data |>
  filter(Experiment_type == "Artificial shading")

# Only intercropping experiments
Intercropping <- IS_Data |>
  filter(Experiment_type == "Intercropping")
```


## Export data as an Excel file
```{r}
# library(openxlsx)
# 
# # Create a new workbook and add a sheet
# wb <- createWorkbook()
# addWorksheet(wb, "Sheet 1")
# 
# # Write data to the sheet
# writeData(wb, "Sheet 1", IS_Data)
# 
# # Save the workbook
# saveWorkbook(wb, "IS_Data_Final.xlsx", overwrite = TRUE)
```





## EDA

#### Univariate: Categorical Variables
```{r}
table(IS_Data$Crop)
table(IS_Data$Crop_type)
table(IS_Data$Experiment_type)
table(IS_Data$Climate_zone)
table(IS_Data$Hardiness_zone)
table(IS_Data$Country)
table(IS_Data$Shade_type)

prop.table(table(IS_Data$Crop_type))
prop.table(table(IS_Data$Experiment_type))
prop.table(table(IS_Data$Climate_zone))
prop.table(table(IS_Data$Hardiness_zone))
prop.table(table(IS_Data$Shade_type))

ggplot(IS_Data, aes(x = Crop)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(IS_Data, aes(x = Crop_type)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(IS_Data, aes(x = Experiment_type)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(IS_Data, aes(x = Climate_zone)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(IS_Data, aes(x = Hardiness_zone)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(IS_Data, aes(x = Country)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(na.omit(IS_Data), aes(x = Shade_type)) + 
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#### Univariate: Numerical Variables
```{r}
summary(IS_Data$Yield)
summary(IS_Data$RSR)
summary(IS_Data$GHI)

ggplot(IS_Data, aes(x = Yield)) + 
  geom_histogram(binwidth = 5)
ggplot(IS_Data, aes(x = RSR)) + 
  geom_histogram(binwidth = 5)
ggplot(IS_Data, aes(x = GHI)) + 
  geom_histogram(binwidth = 100)

ggplot(IS_Data, aes(y = Yield)) + 
  geom_boxplot()
ggplot(IS_Data, aes(y = RSR)) + 
  geom_boxplot()
ggplot(IS_Data, aes(y = GHI)) + 
  geom_boxplot()
```


#### Bivariate - 2 categorical
```{r}
ggplot(IS_Data, aes(x = Crop_type, fill = Climate_zone)) +
  geom_bar(position = "dodge") +
  labs(x = "Crop Type",
       y = "Count",
       fill = "Climate Zone") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(IS_Data, aes(x = Crop_type, fill = Shade_type)) +
  geom_bar(position = "dodge") +
  labs(x = "Crop Type",
       y = "Count",
       fill = "Shade Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Summarize data to compute mean yield per Crop_type and Climate_zone
yield_summary <- IS_Data %>%
  group_by(Crop_type, Climate_zone) %>%
  summarise(Average_Yield = mean(Yield, na.rm = TRUE), .groups = "drop")

# Create bar chart with average yield
ggplot(yield_summary, aes(x = Crop_type, y = Average_Yield, fill = Climate_zone)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Crop Type",
       y = "Average Yield",
       fill = "Climate Zone") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Summarize data to compute mean yield per Crop_type and Hardiness_zone
yield_summary <- IS_Data %>%
  group_by(Crop_type, Hardiness_zone) %>%
  summarise(Average_Yield = mean(Yield, na.rm = TRUE), .groups = "drop")

# Create bar chart with average yield
ggplot(yield_summary, aes(x = Crop_type, y = Average_Yield, fill = Hardiness_zone)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Crop Type",
       y = "Average Yield",
       fill = "Plant Hardiness Zone") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


#### Bivariate - Categorical and Numerical
```{r}
ggplot(IS_Data, aes(x = Crop, y = Yield)) + 
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(IS_Data, aes(x = Crop_type, y = Yield)) + 
  geom_boxplot() +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  xlab("Crop Type") +
  ylab("Yield (%)")

ggplot(IS_Data, aes(x = Climate_zone, y = Yield)) + 
  geom_boxplot()

ggplot(IS_Data, aes(x = Hardiness_zone, y = Yield)) + 
  geom_boxplot()

ggplot(IS_Data, aes(x = Shade_type, y = Yield)) + 
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_boxplot() +
  xlab("Shade Type") +
  ylab("Yield (%)")

ggplot(IS_Data, aes(x = Shade_type, y = RSR)) + 
  geom_boxplot()

ggplot(IS_Data, aes(x = Shade_type, y = GHI)) + 
  geom_boxplot() 

ggplot(IS_Data, aes(x = Experiment_type, y = Yield)) + 
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_boxplot() +
  xlab("Experiment Type") +
  ylab("Yield (%)")
```




#### Boxplots for each crop within a given crop type
```{r}
# Berries
ggplot(filter(IS_Data, Crop_type == "Berries"), aes(x = Crop, y = Yield)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_boxplot() +
  ggtitle("Berries") +
  labs(x = "Crop",
       y = "Yield (%)") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))

# Fruits
ggplot(filter(IS_Data, Crop_type == "Fruits"), aes(x = Crop, y = Yield)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  geom_boxplot() +
  ggtitle("Fruits") +
  labs(x = "Crop",
       y = "Yield (%)") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))

# Fruity vegetables
ggplot(filter(IS_Data, Crop_type == "Fruity vegetables"), aes(x = Crop, y = Yield)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  ggtitle("Fruity vegetables") +
  geom_boxplot() +
  labs(x = "Crop",
       y = "Yield (%)") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))

# Grain legumes
ggplot(filter(IS_Data, Crop_type == "Grain legumes"), aes(x = Crop, y = Yield)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  ggtitle("Grain legumes") +
  geom_boxplot() +
  labs(x = "Crop",
       y = "Yield (%)") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))

# Tubers/root crops
ggplot(filter(IS_Data, Crop_type == "Tubers/root crops"), aes(x = Crop, y = Yield)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black", linewidth = 0.4) +
  ggtitle("Tubers/root crops") +
  geom_boxplot() +
  labs(x = "Crop",
       y = "Yield (%)") +
  #theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))
```

#### Scatterplot grid for Yield and RSR relationship across crop types
```{r warning=FALSE}
ggplot(IS_Data, aes(x = RSR, y = Yield, group = Crop_type, color = Crop_type)) +
  coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_smooth(method = "lm", color = "black", linewidth = 0.5) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Crop_type, scale = "free_y") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")
```
*Note (when shape = Experiment Type): Artificial shading experiments represented by cirle, intercropping experiments represented by triangle


#### Scatterplot grid for Yield and RSR relationship across climate zones for one crop type at a time
```{r}
#Berries
IS_Data |>
  filter(Crop_type == "Berries") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Berries Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#C3 cereals 
IS_Data |>
  filter(Crop_type == "C3 cereals") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("C3 Cereals Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Forages
IS_Data |>
  filter(Crop_type == "Forages") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Forages Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Fruits
IS_Data |>
  filter(Crop_type == "Fruits") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Fruits Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Fruity vegetables
IS_Data |>
  filter(Crop_type == "Fruity vegetables") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Fruity Vegetables Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Grain legumes
IS_Data |>
  filter(Crop_type == "Grain legumes") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Grain Legumes Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Leafy vegetables
IS_Data |>
  filter(Crop_type == "Leafy vegetables") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Leafy Vegetables Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Maize
IS_Data |>
  filter(Crop_type == "Maize") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Maize Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")

#Tubers/root crops
IS_Data |>
  filter(Crop_type == "Tubers/root crops") |>
  ggplot(aes(x = RSR, y = Yield, group = Climate_zone, color = Climate_zone, shape = Experiment_type)) +
  #coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Climate_zone, scale = "free_y") +
  ggtitle("Tubers/Root Crops Yield Response to RSR Across Climate Zones") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")
```

#### Scatterplot grid for Yield and RSR relationship for each shade type
```{r}
ggplot(IS_Data, aes(x = RSR, y = Yield, group = Shade_type, color = Shade_type)) +
  coord_cartesian(ylim = c(0, 300)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  #geom_smooth(method = "loess", color = "black", linewidth = 0.5) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Shade_type, scale = "free_y") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")
```


#### Scatterplot grid for Yield and RSR relationship for each experiment type
```{r}
ggplot(IS_Data, aes(x = RSR, y = Yield, group = Experiment_type, color = Experiment_type)) +
  coord_cartesian(ylim = c(0, 300)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  #geom_smooth(method = "loess", color = "black", linewidth = 0.5) +
  geom_point(alpha = 0.8) +
  theme(legend.position = "none") +
  facet_wrap(~ Experiment_type, scale = "free_y") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")
```


#### Bivariate - 2 numerical
```{r}
IS_Data |>
  ggplot(aes(x = RSR, y = Yield)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", linewidth = 0.75) +
  labs(x = "Reduction in Solar Radiation (%)", y = "Yield (%)")

# IS_Data |>
#   ggplot(aes(x = GHI, y = Yield)) +
#   geom_point() +
#   geom_smooth(method = "loess", color = "blue", linewidth = 0.75) +
#   labs(x = "GHI", y = "Yield (%)")
```







#### Correlations
```{r}
# Compute the correlation matrix
cor_matrix <- cor(numeric_vars)

# Print the correlation matrix
print(cor_matrix)

# Create a heatmap using the base R heatmap function
heatmap(cor_matrix, main = "Correlation Matrix Heatmap")
```
Yield has strongest correlation with RSR, weaker correlation with GHI.

```{r}
# Continuous variables
continuous_vars <- c("Yield", "RSR", "GHI")

# Calculate Pearson correlations between continuous variables
cor_continuous <- cor(IS_Data[, continuous_vars], method = "pearson")
print("Continuous Variable Correlation Matrix:")
print(cor_continuous)

# Categorical variables
categorical_vars <- c("Crop_type", "Experiment_type", "Climate_zone", "Hardiness_zone", "Shade_type")

# For categorical-categorical correlations, we will use chi-squared tests and Cramer's V
library(vcd)  # For Cramér's V

# Loop through all pairs of categorical variables to perform Chi-squared tests and calculate Cramer's V
cat("Categorical-Categorical Correlations (Cramér's V):\n")
for (var1 in categorical_vars) {
  for (var2 in categorical_vars) {
    if (var1 != var2) {
      # Create a contingency table for the two categorical variables
      table_data <- table(IS_Data[[var1]], IS_Data[[var2]])
      
      # Perform Chi-squared test
      chi_square <- chisq.test(table_data)
      cramer_v <- assocstats(table_data)$cramer  # Cramér's V
      cat(paste(var1, "vs", var2, "Chi-squared p-value:", chi_square$p.value, "Cramér's V:", cramer_v, "\n"))
    }
  }
}

# For continuous-categorical correlations, use Point-Biserial correlation (for each continuous and categorical pair)
cat("\nContinuous-Categorical Correlations (Point-Biserial):\n")
for (cont_var in continuous_vars) {
  for (cat_var in categorical_vars) {
    # Convert categorical variable to numeric (as required by cor())
    cat_numeric <- as.numeric(factor(IS_Data[[cat_var]]))
    point_biserial <- cor(IS_Data[[cont_var]], cat_numeric, method = "pearson")
    cat(paste(cont_var, "vs", cat_var, "Point-Biserial Correlation:", point_biserial, "\n"))
  }
}
```

```{r}
# List of categorical variables in your dataset
categorical_vars <- c("Crop_type", "Experiment_type", "Climate_zone", "Hardiness_zone", "Shade_type")

# Loop through all pairs of categorical variables to perform Chi-squared tests or Fisher's Exact Test
cat("Categorical-Categorical Correlations (Cramér's V or Fisher's Test):\n")
for (var1 in categorical_vars) {
  for (var2 in categorical_vars) {
    if (var1 != var2) {
      # Create a contingency table for the two categorical variables
      table_data <- table(IS_Data[[var1]], IS_Data[[var2]])
      
      # Check if the expected counts are too small (less than 5)
      if (any(chisq.test(table_data)$expected < 5)) {
        # Use Fisher's Exact Test with simulation for large or sparse tables
        fisher_result <- fisher.test(table_data, simulate.p.value = TRUE)
        cramer_v <- NA  # Skip Cramér's V as Fisher's test doesn't provide it
        cat(paste(var1, "vs", var2, "Fisher's Exact Test p-value:", fisher_result$p.value, "\n"))
      } else {
        # Perform Chi-squared test with simulation if expected counts are small
        chi_square <- chisq.test(table_data, simulate.p.value = TRUE)
        cramer_v <- assocstats(table_data)$cramer  # Cramér's V
        cat(paste(var1, "vs", var2, "Chi-squared p-value:", chi_square$p.value, "Cramér's V:", cramer_v, "\n"))
      }
    }
  }
}
```









## Regression

```{r}
fullmodel <- lm(Yield ~ ., data = IS_Data)
summary(fullmodel)
```

```{r}
model2 <- lm(Yield ~ Crop_type + RSR + Experiment_type + Climate_zone + Hardiness_zone + GHI + Shade_type, data = IS_Data)

summary(model2)
```

```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))
plot(model2)
```

```{r}
# Perform stepwise selection
final_stepwiseModel <- step(model2, direction = "both", trace = 1)

summary(final_stepwiseModel)
# Crop type, RSR, Experiment type, Climate zone, Hardiness zone
```

```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))
plot(final_stepwiseModel)
```






#### Recreating model from meta-analysis
```{r}
metaCopy <- lm(log(Yield) ~ RSR + I(RSR^2) + RSR*Crop_type + RSR*Experiment_type + RSR*Climate_zone + RSR*Shade_type + RSR*GHI + RSR*GHI*Crop_type + I(RSR^2)*Crop_type + I(RSR^2)*Experiment_type + I(RSR^2)*Climate_zone + I(RSR^2)*Shade_type + I(RSR^2)*GHI, data = IS_Data)

summary(metaCopy)
```

#### Final model from meta-analysis
```{r}
metaFinal <- lm(Yield ~ RSR + I(RSR^2) + RSR*Crop_type, data = IS_Data)

summary(metaFinal)
```

```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))
plot(metaFinal)
```

```{r}
metaFinalLog <- lm(log(Yield) ~ RSR + I(RSR^2) + RSR*Crop_type, data = IS_Data)

summary(metaFinalLog)
```
Log transformation increased R^2 from 0.4601 to 0.5875

Check AIC:
```{r}
AIC(metaFinalLog)
```

Get formula for the model:
```{r}
cat("Model Formula:\n")
formula(metaFinalLog)
```

Diagnostic plots (meta-analysis model):
```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))
plot(metaFinalLog)
```






#### Full model with interactions and quadratic RSR term for stepwise selection
```{r}
full <- lm(Yield ~ RSR + I(RSR^2) + Crop_type + Experiment_type + Climate_zone + Hardiness_zone + Shade_type + GHI + RSR*Crop_type + RSR*Experiment_type + RSR*Climate_zone + RSR*Hardiness_zone + RSR*Shade_type + RSR*GHI + RSR*GHI*Crop_type + I(RSR^2)*Crop_type + I(RSR^2)*Experiment_type + I(RSR^2)*Climate_zone + I(RSR^2)*Hardiness_zone + I(RSR^2)*Shade_type + I(RSR^2)*GHI, data = IS_Data)

summary(full)

#Null model:
null <- lm(Yield ~ 1, data = IS_Data)
```
Full model R^2 = 0.7006


Stepwise selection based on AIC
```{r}
stepAIC <- step(null, scope = list(lower = null, upper = full), direction = "both")

summary(stepAIC)

AIC(stepAIC)
BIC(stepAIC)
```
R^2 = 0.5585, AIC = 3959.877, Residual standard error: 23.04 on 365 df

```{r}
#Get formula for the model
cat("Model Formula:\n")
formula(stepAIC)
```

Stepwise selection based on BIC
```{r}
stepBIC <- step(null, scope = list(lower = null, upper = full), direction = "both", k = log(nrow(IS_Data)))

summary(stepBIC)

AIC(stepBIC)
BIC(stepBIC)
```
R^2 = 0.4441, BIC = 4095.372

```{r}
#Get formula for the model
cat("Model Formula:\n")
formula(stepBIC)
```

Diagnostic plots before sqrt transformation:
```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))
plot(stepBIC)
```










#### Box-Cox to find best transformation
```{r}
# Using final model after stepwise BIC without any transformation yet

library(MASS)
boxcox(stepBIC, lambda = seq(-2, 2, 0.1))
```
Box-cox shows that a square root transformation would be best


#### Square root transformation model
```{r}
final <- lm(sqrt(Yield) ~ I(RSR^2) + Crop_type + I(RSR^2):Crop_type, data = IS_Data)

summary(final)

AIC(final)
BIC(final)
```
R^2 = 0.5322, BIC = 1595.054


Add just climate zone and check R^2 and BIC
```{r}
final <- lm(sqrt(Yield) ~ I(RSR^2) + Crop_type + I(RSR^2):Crop_type + Climate_zone, data = IS_Data)

summary(final)

AIC(final)
BIC(final)
```
R^2 = 0.5552, BIC = 1633.508


Add just hardiness zone and check R^2 and BIC
```{r}
final <- lm(sqrt(Yield) ~ I(RSR^2) + Crop_type + I(RSR^2):Crop_type + Hardiness_zone, data = IS_Data)

summary(final)

AIC(final)
BIC(final)

formula(final)
```
R^2 = 0.5735, BIC = 1605.597

Plant hardiness zone is better predictor than climate zone


Confidence intervals
```{r}
confint(final)
```


```{r}
formula(final)
```


Diagnostic plots: final model
```{r, fig.height=8, fig.width=8}
par(mfrow=c(2,2))
plot(final)
```


VIF
```{r}
library(car)
vif(final)
vif(final, type = "predictor")
```

Centering RSR^2 - did not improve VIF
```{r}
# # Center RSR²
# IS_Data$RSR2_centered <- scale(IS_Data$RSR^2, center = TRUE, scale = FALSE)
# 
# # Refit the model with centered RSR²
# model_centered <- lm(sqrt(Yield) ~ RSR2_centered + Crop_type + RSR2_centered:Crop_type + Hardiness_zone, data = IS_Data)
# 
# # Summary and BIC
# summary(model_centered)
# BIC(model_centered)
# 
# # Check VIF again
# library(car)
# vif(model_centered)
```

Standardizing - did not improve VIF
```{r}
# # Standardize RSR^2
# IS_Data$RSR2_scaled <- scale(IS_Data$RSR^2, center = TRUE, scale = TRUE)
# 
# # Refit the model with standardized RSR²
# model_scaled <- lm(sqrt(Yield) ~ RSR2_scaled + Crop_type + RSR2_scaled:Crop_type + Hardiness_zone, data = IS_Data)
# 
# # Summary and BIC
# summary(model_scaled)
# BIC(model_scaled)
# 
# # Check VIF again
# library(car)
# vif(model_scaled)
```














## Inference

#### ANOVA test for difference in mean yield across shade types
```{r}
oneway.test(Yield ~ Shade_type,
  data = IS_Data,
  var.equal = TRUE # assuming equal variances
)
```


#### ANOVA test for difference in response of yield to RSR between experiment types - on final model
```{r}
final_modelInteraction <- lm(Yield ~ Crop + RSR*Experiment_type + Climate_zone + 
    Hardiness_zone, data = IS_Data)

anova(final_stepwiseModel, final_modelInteraction)
```

#### ANOVA test for significant difference in in the response of yield to RSR across shade types
```{r}
# ANOVA example
anova_model <- aov(Yield ~ Shade_type, data = IS_Data)
summary(anova_model)

# ANCOVA example controlling for RSR
ancova_model <- aov(Yield ~ Shade_type + RSR, data = IS_Data)
summary(ancova_model)
```

```{r}
# ANOVA example
anova_model <- aov(Yield ~ RSR, data = IS_Data)
summary(anova_model)

# ANCOVA example controlling for RSR
ancova_model <- aov(Yield ~ Shade_type + RSR, data = IS_Data)
summary(ancova_model)
```

#### ANOVA test for significant difference in response of yield to RSR between experiment types
```{r}
# Direct test:
expType_interactionModel <- lm(Yield ~ RSR * Experiment_type, data = IS_Data)
expType_reducedModel <- lm(Yield ~ RSR + Experiment_type, data = IS_Data)
anova(expType_reducedModel, expType_interactionModel)

# More nuanced test:

```

There IS evidence of a significant difference in the response of yield to RSR between intercropping and shading-only experiments.





#### Scatterplots with regression lines (code is here because it follows the creation of the models)
```{r}
# Create a data frame with fitted values, confidence intervals, and prediction intervals
IS_Data$Fitted <- predict(metaFinal)  # Fitted values
IS_Data$CI <- predict(metaFinal, interval = "confidence")[, 2:3]  # 95% Confidence intervals
IS_Data$PI <- predict(metaFinal, interval = "prediction")[, 2:3]  # 95% Prediction intervals

ggplot(IS_Data, aes(x = RSR, y = Yield, group = Crop_type, color = Crop_type, shape = Experiment_type)) +
  coord_cartesian(ylim = c(0, 200)) +
  geom_hline(yintercept = 100, color = "black", linewidth = 0.25) +
  
  # Plot black regression line using fitted values
  geom_line(aes(y = Fitted), color = "black", linewidth = 0.5) +
  
  # Plot dark grey confidence intervals
  geom_ribbon(aes(ymin = CI[,1], ymax = CI[,2]), fill = "darkgrey", alpha = 0.4) +
  
  # Plot light grey prediction intervals
  geom_ribbon(aes(ymin = PI[,1], ymax = PI[,2]), fill = "lightgrey", alpha = 0.2) +
  
  geom_point(alpha = 0.6) +
  theme(legend.position = "none") +
  facet_wrap(~ Crop_type, scale = "free_y") +
  xlab("Reduction in Solar Radiation (%)") +
  ylab("Yield (%)")
```


